---
title: "Regression solution"
author: "Oleksii Yehorchenkov"
date: "09 03 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

This is assignment is inspired by practices of [Ryan Miller] (https://remiller1450.github.io/).

## Using regression for predicting binary outcome.

The [Donner Party] (https://en.wikipedia.org/wiki/Donner_Party) dataset documents the fate of members of the Donner Party, a group of 45 pioneers whose migration to California was delayed by a series of mishaps which resulted in the group being stranded in the Sierra Nevada mountains.

**Problem**
Predict survival status members of the Donner Party. The target variable is `Survived`.

## Task

Read the data. Pay attention that data has NA values. It's better to convert target variable to factor.
```{r}
# code for reading data
donner_party <- read.csv("./donner_party.csv", sep=",")
donner_party <- na.omit(donner_party)
donner_party$Survived <- as.factor(donner_party$Survived)
head(donner_party)
```

For prediction we will use only variables `Sex` and `Age`. 

For predicting binary outcome we will use **Generalized Linear Model** (`method = "glm"`) and **caret package**. 

```{r}
# code for regression

library("caret")
training <- train(Survived ~ Sex + Age, data=donner_party, method="glm")
predict <-predict(training, newdata=donner_party)

```

What is your in-sample accuracy? Build a confusion matrix

```{r}
# code for confusion matrix

confusion_matrix = confusionMatrix(predict, donner_party$Survived)
confusion_matrix

```
Use a cross-validation with `method = "repeatedcv"`, `number = 5`, `repeats = 10`. Does cross-validation improve the accuracy?

```{r}
# code for cross-validation

train_control <- trainControl(method = "repeatedcv", number = 5, repeats = 10)
training2 <- train(Survived ~ Sex + Age, data = donner_party, method = 'glm', trControl = train_control)
predict2 <- predict(training2, newdata = donner_party)

confusion_matrix2 = confusionMatrix(predict, donner_party$Survived)
confusion_matrix2

```
** With cross-validation the accuracy remains the same **
